---
title: "esDataViz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## esData loading

Before we do any data visualization. Let's first remove some redundant columns from the original dataset.
```{r removecolumns, include=FALSE}
# load data into tibble
load("D:/1VDR/emotion-project/esDataF.Rdata")
es_tibble <- as_tibble(esDataF)

# remove columns
es_subset <- es_tibble %>% 
  select(-ends_with(".low"), -ends_with(".high"))
```

And we can take a look at the summary of the remaining dataset:
```{r datasummary}
summary(es_subset)
```

## Analyze thoughts

The column mentionedTotal shows the total number of thoughts a user report at one time. Let's see the histogram and boxplot of this column.

```{r pressure, echo=FALSE}
ggplot(data=es_subset) + 
  geom_histogram(mapping=aes(x=mentionedTotal), fill="blue", boundary=0, binwidth=1) + 
  scale_y_log10(name="Number of Rows")

ggplot(data=es_subset) + geom_boxplot(mapping=aes(x=mentionedTotal), fill="blue")
```

From the histogram and boxplot, we can tell a lot of records have 0 thoughts, so let's filter them out:
```{r filterzero}
es_subset <- es_subset %>% 
  filter(mentionedTotal!=0)
```

Then, let's sum up each thoughts and visualize which thought gets more attentions:
```{r sumthought, include=FALSE, error=TRUE}
thoughts <- es_subset %>%
  select(tFinances,tVices,tWondering,tWorkSchool,tFuture,tHousework,tMundane,tMyHelathWellb,tPast,tPasttimes,tPolReligion,tRegretDisIn,tSocial) %>%
  colSums(na.rm = TRUE)
thoughts <- sort(thoughts, decreasing = TRUE)
#c("tFinances","tVices","tWondering","tWorkSchool","tFuture","tHousework","tMundane","tMyHelathWellb","tPast","tPasttimes","tPolReligion","tRegretDisIn","tSocial")
```

```{r visualizesum}
ggplot(mapping = aes(x = reorder(names(thoughts), -unname(thoughts)), y = unname(thoughts))) +
  geom_col(fill="blue") +
  scale_y_continuous(name = "Number of Rows") +
  scale_x_discrete(name = "Thoughts") + 
  theme(axis.text.x = element_text(angle = 90))
```

We can also sum up each domain and visualize the top ten domains and the bottom ten domains:
```{r sumdomain, include=FALSE, error=TRUE}
domains <- es_subset %>%
  select(dAccomplish,dBills,dBook,dCantAfford,dCarGoals,dCommute,dDeath,dDeceased,dDislike,dDIY,dDream,dDrugDrink,dEducateKids,dErrands,dExer,dExPartner,dFantasy,dFlawedAppearance,dFood,dFrust,dFunny,dGetJob,dHappyPastEvents,dHobby,dHowMuchLove,dIdeate,dImageSocial,dImplausibleNeg,dImplausiblePos,dInterwork,dOthersProbs,dOthersSoc,dpainfulEvent,dPastPeriod,dPersonalGoals,dPlausibleNeg,dPlausiblePos,dPoliticsNewSocial,dPosEvent,dRehearseLearn,dRelationReg,dReli,dRomantic,dSchedule,dSecondG,dSelfToDo,dSex,dShopping,dSleep,dSportPlay,dStartB,dThinkOfMe,dTvMovie,dUpwardCounter,dVideo,dWatchSport,dWeather,dIssuesWork,dLeisure,dLikeAppear,dLovedOneWellH,dMakeMoney,dMeaning,dMistakes,dMusic,dMyArt,dMyHealth,dNegEvent) %>%
  colSums(na.rm = TRUE)
domains <- sort(domains, decreasing = TRUE)
domaintop10 <- head(domains, n=10)
domainbottom10 <- tail(domains, n=10)
```
```{r visualizesum2}
ggplot(mapping = aes(x = reorder(names(domaintop10), -unname(domaintop10)), y = unname(domaintop10))) +
  geom_col(fill="blue") +
  scale_y_continuous(name = "Number of Rows") +
  scale_x_discrete(name = "Top 10 domains") + 
  theme(axis.text.x = element_text(angle = 90))

ggplot(mapping = aes(x = reorder(names(domainbottom10), -unname(domainbottom10)), y = unname(domainbottom10))) +
  geom_col(fill="blue") +
  scale_y_continuous(name = "Number of Rows") +
  scale_x_discrete(name = "Bottom 10 domains") + 
  theme(axis.text.x = element_text(angle = 90))
```
